HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-09 paper/main.tex
fff

  Semantic evaluation of Machine Translation (MT) systems, informative as for
  semantic discrepencies between the translation and the source text,
  has become a necessity, given recent interest in semantics-aware MT systems.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output, whose interpretation is often unclear. 
  We experiment with four language pairs, demonstrating the broad applicability of
  the measure, as well as its reliability, reflected by the inter-annotator agreement
  scores.

@@

Human judgement is the central, and perhaps ultimate, criterion for estimating the
quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the entire sentence, as is reflected
by the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.

These problems were partially addressed by measures that decompose over sub-parts of the sentence,
most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
translation. More recent work on MT evaluation, most notably the MEANT measure \cite{lo2011structured},
proposed measures that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of their verb argument structures.
While allowing a localization of the semantic errors in the translation,
focusing on verb argument structures misses out on many very pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus provides only a limited
perspective on the semantic similarity between the output and reference.
Moreover, assigning a semantic structure to MT output, which is often low in quality,
is difficult and arguably ill-defined.

@@

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires the semantic annotation of the source sentence, thus avoiding the 
annotatation of machine-generated text, a difficult if not ill-defined task, and allowing the re-use
of the source semantic annotation for measuring the quality of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
See \secref{sec:hmeant_comp} for further discussion.

We conduct experiments with four language pairs, English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreements which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's agreement does not diminish with the sentence length.

While human evaluation is not a scalable solution,
it provides the foundations and justification for defining a measure.
The paper thus focuses on human evaluation, but concludes
by outlining ongoing work to construct semi- and fully-automatic variants of HUME.

@@

\paragraph{Machine Translation Evaluation.}
MT evaluation measures has been traditionally pursued through human-based
evaluation that quantifies the adequacy of the translation in preserving the meaning 
of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is still accepted as the standard evaluation measure, and is
estimated either in terms of the relative ranking of the outputs of multiple systems
as is standard in the WMT shared tasks, e.g., \cite{bojar2015findings}, or by assigning
indiviaul adequacy/fluency scores for each sentence, a procedure recently improved
by \newcite{graham2015improving}.

While providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
which has motivated the construction of cheaper methods that correlate with human
judgement.

Scalability has been addressed by scores that decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. The most commonly used measures decompose
over n-grams or even individual words, with examples such as
the BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005}.
Other methods employ a string edit distance for determining the similarity between the
reference and translation \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
to evaluate experiments no a massive scale, they 
decompose over the words and n-grams of the sentence,
and are thus better suited for localizing errors
when evaluating shallow MT methods, and are less telling for identifying the semantic
nature of translation errors in the output.

@@

To address this shortcoming, more structure-aware methods were proposed, with the 
purpose of quantifying how different the reference and translation output are in terms 
of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
quantifying the similarity between the syntactic structures of the reference and translation.
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed an evaluation measure that incorapotes features
that capture the overlap between the output and reference in terms of their
semantic (argument structure phenomena), lexical and morphosyntactic features.

Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT and
its human variant HMEANT, which compares the reference and translation in terms of the similarity
of their semantic role labeling structure. HMEANT shares much of HUME's motivation for providing a
measure that decomposes over semantic units, rather than strings, but focuses on verbal
argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
METEOR measures, in not requiring a reference, as it compares the source directly with the output
translation. We see this difference as crucial where semantic analysis is concerned, as it
is often impossible to assign a semantic structure to a low quality MT output sentence.
See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
and HMEANT.

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units} , are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear one
or more categories, indicating the role of the sub-unit in the relation the unit represents. UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.\oa{add figure}

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of distinctions definitions,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2012}. This is an indication that the scheme reflect a common layer of representation that
should be preserved in translation, and so should be represented in the translation output too.
We note that that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure. For example, ``after he graduated, Tom moved to NYC'' conveys a similar meaning
to ``after graduation, Tom moved to NYC''.

While most broad-coverage semantic work has focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.

Other approaches have been proposed for semantic representations, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representations,
these approaches require linguistic expertise for their annotatation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable and
easily interpretable means for our purposes.
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

@@

\subsection{Annotation Guidelines}\label{sec:guidelines}
In this section we summarise the procedure for annotating machine translation output.

@@

\subsection{Annotation Process}

For the annotation of the machine translation, we recruited seven annotators, two for each of the target languages apart from Czech. They were
all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
to annotate small numbers of sentences each for the purpose of calculating inter-annotator agreement. In \tabref{tab:annot}
we show the total number of sentences and nodes annotated by each annotator, where we consider the additional Czech
annotators as a single annotator.

\emph{Show numbers of annotations, and timings}

\emph{Describe the UCCA annotation}

>>>>

  Recent interest in semantics-aware MT systems underscores te importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of structure-aware MT, much in the same way that string-based measures,
  such as BLEU, support advances with shallower MT systems.  
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating the broad applicability of
  the measure, as well as its reliability, reflected in the inter-annotator agreement
  scores.

@@

Human judgement is the central, and perhaps most fundamental, criterion for
estimating the quality of an MT system.
Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
or the relative ranking of several possible translations, are problematic in two ways.
First, as the quality of translation is determined by multiple factors, it is difficult
to assign a single number that quantifies the quality of the whole entire sentence. This
is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
as a function of the sentence length \cite{Bojar:2011}.
Second, a sentence-level quality score does not indicate what parts of the sentence
are erroneously translated, and hence cannot inform developers in overcoming these errors.

These problems were partially addressed by measures that decompose over sub-parts of the sentence,
most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
translation. More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
MEANT \cite{lo2011structured}).
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence.
Importantly, HUME only requires semantically annotating the source sentence,
thus avoiding the annotatation of machine-generated text 
and allowing re-use of the source semantic annotation for measuring the quality
of different translations of the same source sentence.
By basing the measure on a direct comparison
of the source and translation we avoid the difficulties engendered by using
reference translations, albeit at the cost of imposing the additional requirement
that the evaluator be proficient in both the source and target languages.
See \secref{sec:hmeant_comp} for further discussion.

We conduct experiments with four language pairs, English-German, English-Polish,
English-Romanian and English-Czech, and report efficiency (time of annotation) and
inter-annotator agreement scores, as well as qualitative comparison of the resulting
HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
agreements which are on par with those of standard sentence-level ranking measures,
but unlike them, HUME's agreement does not diminish with the sentence length.
The paper concludes by outlining ongoing work to construct semi- and fully-automatic variants of HUME.

@@

\paragraph{Machine Translation Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is the standard measure, and is
estimated either in terms of the relative ranking of the outputs of multiple systems
e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
indiviaul adequacy/fluency scores to each sentence, a procedure recently improved
by \newcite{graham2015improving}.
However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
which has motivated the construction of cheaper methods that correlate with human
judgement.

Scalability is addressed by scores that decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. The most commonly used measures decompose
over n-grams or even individual words, e.g., 
the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
Other methods employ a string edit distance for determining the similarity between the
reference and translation \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
to evaluate experiments no a massive scale, they 
decompose over the words and n-grams of the sentence,
and are thus better suited for localizing errors
in the output of shallow MT methods, and are less telling for identifying the semantic
nature of translation errors in the output.

@@

To address this shortcoming, more more recent work quantifies
the similarity of the reference and translation output in terms
of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
quantifying the similarity between the syntactic structures of the reference and translation.
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed an evaluation measure that incorapotes features
that capture the overlap between the output and reference in terms of their
semantic (argument structure phenomena), lexical and morphosyntactic features.

Perhaps the most notable attempt at semantic MT evaluation is MEANT and
its human variant HMEANT, which quantifies the similarity between
the reference and translation in terms of the overlap in
their verbal argument structures structures, and associated semantic roles.
HMEANT shares much of HUME's motivation for providing a
measure that decomposes over semantic units, rather than strings, but focuses on verbal
argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
METEOR measures, in not requiring a reference. Instead,
it compares the source directly with the output
translation. We see this difference as crucial where semantic analysis is concerned, as it
is often impossible to assign a semantic structure to a low quality MT output sentence.
See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
and HMEANT.

\begin{figure}

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
    \node (ROOT) [fill=black, circle] {}
    child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
    child {node (H1) [fill=black,circle] {}
      child {node {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [fill=black,circle] {}
      child {node {} edge from parent[draw=none]}
      child {node {} edge from parent[draw=none]}
      child {node (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node {moved} edge from parent node[right] {$P$}}
      child {node [fill=black,circle] {}
        child {node {to} edge from parent node[right] {$R$}}
        child {node {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation.}
\end{figure}

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear
a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.
See \figref{fig:ucca_example}.

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of distinctions definitions,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2015conceptual}. This is an indication that the scheme reflects
a common layer of representation that in a correct translation
is mostly shared between the translation and the source. 
We note that that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure, and are thus less motivated as a basis for a semantic measure.
For example, ``after he graduated, Tom moved to NYC''
conveys a similar meaning to ``after graduation, Tom moved to NYC''.

While most broad-coverage semantic work focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.

Other approaches have been proposed for semantic representation, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
these approaches require linguistic expertise for their annotatation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable for basing
an interpretable MT evaluation on. 
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

@@

\subsection{Annotation Guidelines}\label{sec:guidelines}

@@

\subsection{Annotation Process}
\begin{itemize}
  \item Numbers of annotators, IDs
  \item Length of annotation, timings
  \item Sentences annotated
\end{itemize}

For the annotation process, we recruited seven annotators, two for each of the target languages apart from Czech. They were
all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
to annotate small numbers of sentences each for the purpose of calculating inter-annotator agreement. In \tabref{tab:annot}
we show the total number of sentences and nodes annotated by each annotator, where we consider the additional Czech
annotators as a single annotator.
